### 2025-12-05 Paper/paper_2025.tex

    \section{Data Construction}

@@

    \import{Objects}{boardand_people.tex}

    I create a list of overlapping board member candidates by identifying any names that occur on multiple hospital boards in the same year. However, it is very common for hospitals within a system to largely share the same board. This is not the type of variation I seek to investigate in this paper, since hospitals within systems are already heavily affiliated. Therefore, I define the relevant type of overlap as someone being on the board of multiple hospitals that are otherwise \textit{unaffiliated}. Hospitals within systems are still eligible to have overlap with other hospitals that do not belong to the same system as they do. Further, In the hospital merger literature, mergers have been found to be more significant when the two hospitals involved are located geographically close (\cite{cooper2019price}). Therefore, I define two types of overlap: overlap with an otherwise non-affiliated hospital in the same market, or overlap with an otherwise unaffiliated hospital in a different market. All other hospitals have no board overlap with otherwise unaffiliated hospitals. I use the market definitions developed by Dartmouth Atlas, Hospital Referral Regions (HRRs), which are meant to capture geographic regions of the average patient's choice set of hospitals.

    The final sample limitation I make for simplicity is to remove hospitals that gain and then lose overlapping board members in the sample. Since treatment will be turning on overlapping board member, I remove those who "lose" their treatment.

@@

    The full sample of hospitals

    \subsection{Summary Statistics}

    First, I present a summary statistics table showing characteristics of all hospitals in the final sample, which consists of 1,467 nonprofit hospitals, almost all of which are adult general hospitals. They have, on average, 178 beds and 14 board members. Almost 40\% of the hospitals in the sample are academic medical centers, almost 70\% of the sample are located in metropolitan areas, and 56\% belong to a system.

    %\import{Objects}{all_summarystats.tex}

    Eight percent of hospitals in the sample have overlapping board members with an otherwise unaffiliated hospital. In Figure \ref{fig:connected_percent}, I show that this percentage remains relatively constant over time, with a slight decline from 129 affiliated hospitals in 2017 to 114 affiliated hospitals in 2021. I include 2022 values, but there are less total firms in the sample due to lag in tax forms being made available publicly.

    \begin{figure}[ht!]
        \centering
        \vspace{9mm}
        \caption{Percent of hospitals that have a board affiliation}
        \includegraphics[width=.8\textwidth]{Objects/connected_percent.pdf}
        \label{fig:connected_percent}
    \end{figure}

    As mentioned previously, common board members should be more likely to affect behavior within a hospital market, where hospitals compete for the same patient population. In Figure \ref{fig:connected_HRR_percent}, I show the prevalence of having shared board members within a hospital market. Of the 306 HRRs in the US, approximately 17\% have at least one pair of board affiliated hospitals within the market. This proportion decreases slightly over time, with 15\% of HRRs having connected hospitals in 2021, a nominal decrease of 8 markets losing hospitals with affiliation. This is likely due to increased system affiliations, which mechanically decreases board overlap since I do not include hospitals in the same system as eligible for having board overlap.

    \begin{figure}[ht!]
        \centering
        \caption{Percent of HRRs containing board affiliated hospitals}
        \includegraphics[width=.8\textwidth]{Objects/connected_HRR_percent.pdf}
        \label{fig:connected_HRR_percent}
    \end{figure}

    Next, I present the geographic distribution of these pairs in Figure \ref{fig:connected_maps}. The blue lines represent hospitals that share a common board member, and the red dots are all other hospitals. Connected hospitals are distributed fairly evenly across the US, apart from relatively few connections on the west coast relative to the population. Over time, the connections remain fairly consistent geographically. Taken together, these show that board overlap is a prevalent occurrence among nonprofit hospitals in the US from 2017 onward.

    \begin{figure}[ht!]
        \centering
        \caption{Geographic distribution of board affiliated hospitals over time}
        \includegraphics[width=\textwidth]{Objects/connected_maps.pdf}
        \label{fig:connected_maps}
    \end{figure}

    I now shift to examining characteristics of hospitals that share board members. I consider pairs of connected hospitals and characteristics of the pair in Table \ref{tab:hospital_pair_types}. Each hospital in a pair is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or belonging to a system, all characteristics in the AHA data. Additionally, the median bed size of all hospitals in the sample is 103, and I define hospitals as being either small or large depending on whether they fall below or above this median.

@@

    Two-thirds of board affiliated pairs are between two adult general hospitals, and 10\% of pairs are between a specialty and a general hospital. Thus, for the majority of hospitals that have board interlock, the hospital they share a board member with is a direct competitor. Additionally, there is significant variation in the ownership type of hospitals within affiliated pairs. Forty-five percent of pairs are between two non-system affiliated hospitals, and 36\% of pairs are between an independent hospital affiliated with a hospital belonging to a system. The lowest proportion of shared board member pairs is made up of two hospitals belonging to different systems.

    There is also variation in whether the hospitals are the same size or not. Half of the hospital pairs consist of one small and one large hospital, and the other half are split evenly between two small or two large hospitals. However, the average difference in the number of beds of the two hospitals in a pair is very large: an average difference of 254 beds. This suggests that for hospitals of opposite size, the difference is drastic. The average distance between hospitals that share common board members is 35 miles, or 57 kilometers. In total, there are 217 hospitals that share a board member with at least one other hospital at some point in the sample.

    \subsection{Correlations with Hospital Behaviors}

    I now merge the hospital board data to other publicly available data sets to examine descriptive differences in the hospital behaviors discussed in Section \ref{sec:hospbehaviors} for hospitals with and without board overlap. The broad categories of hospital behaviors that might be affected by affiliation through shared board members are price/quality, investment decisions, and differentiation, which are discussed in Section \ref{sec:hospbehaviors}. While I do not capture every aspect of hospital behavior, I select a number of variables that fall into these categories to examine.

    First, from the Center for Medicare and Medicaid Services (CMS) Hospital Compare data, I merge information on each hospital's overall star rating. This rating combines measures of quality into one index that ranges from 1 to 5: readmission rates, mortality rates, safety, timeliness of care, effectiveness of care, and patient experience. This is a very aggregated measure, but may capture big-picture differences in quality.

    Next, I include data on different types of investments made into the hospital from the Healthcare Cost Reporting Information System (HCRIS) data. These variables include the dollar amounts spent on health information technology purchases, fixed equipment, and movable equipment.

    There are a plethora of ways one could measure differentiation in hospitals. I focus on two categories of differentiation: service offerings and patient population. To measure different service offerings, I first construct a measure of concentration of beds devoted to categories of services in the AHA survey. For services $j$ in the list of all possible services\footnote{General medical and surgical (adult), General medical and surgical (pediatric), Obstetric care, Medical/surgical intensive care, Cardiac intensive care, Neonatal intensive care, Neonatal intermediate care, Pediatric intensive care beds, Burn care, Other Special Care, Other intensive care, Physical Rehabilitation care, Alcohol/drug abuse or dependency inpatient care, Psychiatric care, Skilled nursing care, Intermediate nursing care, Acute long term care, Other long-term care, Other care}, the concentration of bed allocation is defined as

    $$concentration_{bed} = \sum_{j}\left(\frac{beds_j}{total beds}\right)^2.$$

    \noindent That is, a higher bed concentration indicates more specialization in bed allocation, where a lower concentration indicates more services spread out in bed allocations. Additionally, I include a second concentration measure that relies less on the facility and more on the conditions patients are treated with. The CMS Provider Utilization files contain measures of the percent of patients seen for different conditions. Therefore, for services $k$ in the list of conditions in the CMS data\footnote{Percent Of Beneficiaries Identified With Combined Cancer Flag For 6 Cancer Indicators, Percent Of Beneficiaries Identified With Chronic Obstructive Pulmonary Disease, Percent Of Beneficiaries Identified With Chronic Kidney Disease, Percent Of Beneficiaries Identified With Heart Failure And Non-Ischemic Heart Disease}, I define the concentration of patient conditions in each hospital as

    $$HHI_{patient} = \sum_{k}\left(\frac{patients_k}{total patients}\right)^2.$$

    Finally, I include indicators for whether the hospital has a NICU or Cath Lab, both profitable services for hospitals to offer. I also consider differentiation by the patient population admitted to the hospital. There is information in the HCRIS data about how many discharges are for patients on Medicaid or Medicare. There is also information in the CMS Provider Utilization Files about the average risk score of Medicare patients seen in the hospital. I include all of these variables to capture something about the aggregate patient population being treated in different hospitals.

    In Table \ref{size_characteristics}, I present averages of all of these variables for four categories of hospitals: whether they are small (less than 103 beds) or large (greater than 103 beds), and whether they share board members with another competing hospital. I limit this sample to only general hospitals for relevant comparison.

   \import{Objects}{size_characteristics}

   Small hospitals with and without shared board members have a very similar average star rating of 3.4, while large hospitals with shared board members have a slightly smaller average star rating than large hospitals without board overlap. In all categories of investment, small hospitals are relatively similar, with slightly more investment in health IT and fixed equipment for board affiliated small hospitals. Large hospitals are also generally similar in how they invest in health IT and fixed equipment relative to how much they spend on average, but hospitals with shared board members spend significantly more on movable equipment than hospitals without board overlap.

   There do not seem to be meaningful differences in the concentration of services offered or conditions treated among hospitals with and without board overlap, whether they are small or large. Similarly, there are no clear differences in the patient composition of hospitals with and without board overlap. This is consistent even among the facilities of these hospitals, with no distinct differences in whether hospitals have a NICU or Cath lab based on board overlap.

   Another important level of variation in this sample of hospitals could be whether they belong to a system or not, as this often dictates a number of the behaviors discussed. Therefore, I present a similar table of means for hospitals with and without board overlap that do or do not have a system affiliation in Table \ref{sys_characteristics}.

   \import{Objects}{system_characteristics}

    Among hospitals that do not belong to a system, those without board overlap have a slightly higher quality rating than those without board overlap. However, the opposite is true among hospitals that do belong to a system, where those with board overlap have a higher quality rating. The most significant difference in investment is among hospitals that belong to a system for investment in movable equipment. Those with board overlap invest over 20 million in movable equipment, while those without board overlap only invest 8.3 million. The other comparisons of investment show no drastic differences. In-system and out-of-system hospitals both show higher levels of concentration of services when they do not have board overlap. However, the patient composition is similar within the categories of system affiliation. Independent hospitals that do not have board overlap are more likely to have a Cath lab, but otherwise hospitals are similar in in service offerings.

    While these statistics do not establish a causal link between board overlap and hospital behavior, they highlight key differences across hospital categories. Notably, hospitals with overlapping board members exhibit distinct patterns in quality and investment decisions. These differences underscore the need for further research to explore board overlap as a potential mechanism for affiliation and behavioral changes in health care.

>>>>

    \section{Data and Summary Statistics}

@@

    I present summary statistics of characteristics of the people in the data in Table \ref{tab:board_people}. There are just under 37,000 people in the data, 63\% of which are board members and the remaining 37\% are executives. Board members are slightly moe likely to be medical doctors than executives, but the fraction of doctor board members is still relatively low at 18\%. Nurses are consistent across board and executives, but only 3\% of the sample are nurses. A negligible number of people claim to have a health administration degree. Executives are slightly more likely to be female according to the name algorithms, with 38\% of executives estimated to be female and 32\% of board members estimated to be female. I also include average team composition of hospital boards and executive teams, where these trends largely hold at the hospital level. On average, hospital boards contain 7 members and executive teams contain 10 members.

    \import{Objects}{boardandexec_people.tex}

    I create a list of overlapping board member candidates by identifying any names that occur on multiple hospital boards in the same year. However, it is very common for hospitals within a system to largely share the same board. This is not the type of variation I seek to investigate in this paper, since hospitals within systems are already heavily affiliated. Therefore, I define the relevant type of overlap as someone being on the board of multiple hospitals that are otherwise \textit{unaffiliated}. Hospitals within systems are still eligible to have overlap with other hospitals that do not belong to the same system. I exclude hospitals that gain or lose overlapping board members several times in the sample period.

    In the hospital merger literature, mergers have been found to be more significant when the two hospitals involved are located geographically close (\cite{cooper2019price}). Therefore, I define two types of overlap: overlap with an otherwise non-affiliated hospital in the same market, or overlap with an otherwise unaffiliated hospital in a different market. I use the market definitions developed by Dartmouth Atlas, Hospital Referral Regions (HRRs), which are meant to capture geographic regions of the average patient's choice set of hospitals.

@@

    In Figure \ref{fig:connected_percent} I show the prevalence of overlapping board members or executives over time. In 2017, approximately 5\% of hospitals have an overlapping board members or executive with another hospital in the same market that is not formally affiliated otherwise. That is, these hospital are not under common ownership. This percentage increases over time, with 10\% of hospitals sharing a key leader in their market in 2022. I also graph the percent of HRR markets where hospitals within that market share board members or executives. Approximately 15\% of HRRs are potentially affected by this type of affiliation. Sentence here about putting this into perspective relative to hospital consolidation as a whole.

    \begin{figure}[ht!]
        \centering
        \vspace{9mm}
        \caption{Prevalence of Board/Executive Interlock}
        \includegraphics[width=.9\textwidth]{Objects/connected_percent.pdf}
        \label{fig:connected_percent}
    \end{figure}

    Next, I present the geographic distribution of these pairs in Figure \ref{fig:connected_maps}. The blue lines represent hospitals that share a common board member, and the red dots are all other hospitals. Connected hospitals are distributed fairly evenly across the US, apart from relatively few connections on the west coast relative to the population. Over time, the connections remain fairly consistent geographically.

    \begin{figure}[ht!]
        \centering
        \caption{Geographic Distribution of Board/Executive Interlock}
        \includegraphics[width=\textwidth]{Objects/connected_maps.pdf}
        \label{fig:connected_maps}
    \end{figure}

    Hospitals in the sample are either always connected, never connected, or become connected at some point in time. I show characteristics of hospitals in these categories in Table \ref{tab:hosp_group_stats}.

    \import{Objects}{hosp_group_stats.tex}

    I now present characteristics of pairs of hospitals that are connected by board members or executives in Table \ref{hospital_pair_types}. In the first column, I show pair characteristics of hospitals connected within the same HRR. In the second column, I present characteristics of hospitals connected to otherwise unaffiliated hospitals located in a different HRR. Therefore, all hospital pairs in the table are connected, but only the first column are connected to hospitals that are potentially direct competitors. Each hospital in a pair is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or belonging to a system, all characteristics in the AHA data. Additionally, the median bed size of all hospitals in the sample is 101, and I define hospitals as being either small or large depending on whether they fall below or above this median. There are typically 1.5 connections in the same HRR and almost 6 connections in different HRRs.

@@

    Most pairs, 76-78\% are two general hospitals. Thus, for the majority of hospitals that have board/executive interlock in the same HRR, the hospital they share a leader with is a direct competitor. I also include characteristics of system versus independent ownership. For there is a nearly even spread of the type of ownership of the pair. 42\% of these pairs are made up of one independent hospital and one system-owned hospital, while 37\% are made up of two independent hospitals and the rest are made up of two hospitals that are both system owned but belong to two different systems. These figures are slightly different for connections in different HRRs, where a larger portion of connections, 45\%, are made up of two system-owned hospitals of two different systems.

    In both same and different HRR connected pairs, 50\% of the connections are such that one hospital is small in bed size and one is large in bed size. The average bed difference is over 200, indicating that the size difference of the average hospital connection is drastic. what's the implication of this? Finally, connected hospitals in the same HRR are approximately 62 kilometers, or 38 miles, away from the other.

    add outcome graphs
